Recursive algorithms, branching coefficients and applications

### 2010-07-06 AffineLieAlgebras/InformalNotes/article.tex
Paper is shortened for JPhysA submission, figures are converted to eps and renamed according to IOP LaTeX Guidelines.

Such effects are due to the fact that the subalgebras $\mathfrak{a}$ and $\mathfrak{a}_{\bot}$ must not form a direct sum in $\mathfrak{g}$.  
Consider the case of such a regular embedding $\mathfrak{a}\subset \mathfrak{g}$ where both algebras are simple and the diagram of the subalgebra $\mathfrak{a}_{\bot}$ is not a subdiagram of the extended Dynkin diagram $\mathfrak{g}$.
 Drop the subdiagram of $\mathfrak{a}$ and the node $\alpha'$ that connects it with all the remaining nodes of the diagram of $\mathfrak{g}$. The remaining diagram is the diagram of the algebra $\mathfrak{\bar{a}}$ with $\mathrm{rank}(\mathfrak{\bar{a}}) = \mathrm{rank}(\mathfrak{g})-\mathrm{rank}(\mathfrak{a})$. It is clear that $\mathfrak{\bar{a}}\subset \mathfrak{a}_{\bot}$. So the question is whether $\mathfrak{a}_{\bot}$ has additional roots,  not the roots of $\mathfrak{\bar{a}}$ but the linear combinations of them. The answer is positive  when the set of angles between the roots of $\mathfrak{\bar{a}}$ does not contain all the angles between the roots of $\mathfrak{a}$. Then by reflecting the roots of $\mathfrak{a}$ by $s_{\alpha'}$ we get the additional roots of $\mathfrak{a}_{\bot}$. 
\begin{table}[ph]
\label{tab:diagrams}
\noindent  \centering{
\begin{tabular}{|l|l|l|l|}
  \hline
  $\mathfrak{g}$ & Extended diagram of $\mathfrak{g}$ & Diagrams of the subalgebras $\mathfrak{a},\; \mathfrak{a}_{\bot}$ \\
  \hline
  $A_n$ & \includegraphics{A_series_ext_diag.eps} & \includegraphics{A_series_ext_split.eps} \\
  \hline
  $B_n$ & \includegraphics{B_series_ext_diag.eps} & \includegraphics{B_series_ext_split.eps} \\
  \hline
  $C_n$ & \includegraphics{C_series_ext_diag.eps} & \includegraphics{C_series_ext_split.eps} \\
  \hline
  $D_n$ & \includegraphics{D_series_ext_diag.eps} & \includegraphics{D_series_ext_split.eps} \\
  \hline
\end{tabular}
\caption{Subalgebras $\mathfrak{a},\; \mathfrak{a}_{\bot}$ for the classical series}}
\end{table}
If $\mathfrak{a}=B_{r_{\mathfrak{a}}}$  we see that $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\;\alpha_j,\; 1<i<j\leq r-r_{\mathfrak{a}}\}$ and $\mathfrak{a}_{\bot}=B_{r-r_{\mathfrak{a}}}$. This is the only case where the simple roots of $\mathfrak{a}_{\bot}$ can not be obtained from the extended Dynkin diagram, as can be seen in the Table \ref{tab:diagrams}.

@@

Let's describe the reduction of the fundamental representation of $L^{(1,0)=w_1}_{B_2}$ (the black vector in Figure \ref{fig:B2_A1}).
\begin{figure}[th]
  \noindent\centering{
    \includegraphics[width=80mm]{B2_A1.eps}
  }
  \caption{Regular embedding of $A_1$ into $B_2$}
  \label{fig:B2_A1}
\end{figure}
Circles indicate the weights of the singular element $\Psi^{(w_1)}$.
Now we are to factorise the Weyl group $W$ by $W_{\bot}=\left\{\omega_1\right\}$ and to construct the set $\left\{\omega(\mu+\rho)-\rho,\; \omega\in W_{\bot}\backslash W\right\}$. 
\begin{figure}[t]
  \noindent\centering{
    \includegraphics[width=80mm]{B2_A1_2.eps}
  }
  \caption{Anomalous weights and the corresponding $\mathfrak{a}_{\bot}=A_1$-modules for the embedding $A_1\subset B_2$}
  \label{fig:B2_A1_2}
\end{figure}
On the Figure \ref{fig:B2_A1_2} the weights of  the corresponding $\mathfrak{a}_{\bot}=A_1$-modules $L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}$ are indicated. Projecting them onto the root space of the subalgebra $\mathfrak{a}=A_1$ we get the anomalous weights with the corresponding multiplicities:
\begin{equation}
  \label{eq:25}
  (1,2),\; (0,-3),\; (-4,3),\; (-5,-2)
\end{equation}
For the set $\Gamma$ (using the definition (\ref{phi-d},\ref{fan-defined})) we have
\begin{equation}
  \label{eq:22}
  \Gamma_{A_1\subset B_2}=\left\{ (1,2),\; (2,-1) \right\}.
\end{equation}
Here the second component denotes the value of $s(\gamma)$. Applying this fan inside the $\bar{C}^{(0)}_{\mathfrak{a}}$ we get zeros for the weights greater than the first anomalous vector $(1)$, here $k^{(1,0)}_1=2$. For the last weight in $\bar{C}^{(0)}_{\mathfrak{a}}$ the formula (\ref{recurrent-relation}) gives
\begin{equation}
  \label{eq:23}
  k^{(1,0)}_{0}=-1\cdot k^{(1,0)}_2 +2\cdot k^{(1,0)}_1 - 3\cdot \delta_{0,0} = 1.
\end{equation}
The recurrence property defines the branching.

\subsection{Embedding of $B_2$ into $B_4$}
\label{sec:someth-high-dimens}
Consider the regular embedding $B_2 \longrightarrow B_4$.
The corresponding Dynkin diagrams are presented in the Figure \ref{fig:dynkin}.
\begin{figure}[h]
  \centering
  \includegraphics[width=50mm]{B4_B2_2A1.eps}
  \caption{The regular embedding $B_2\subset B_4$ described by dropping the node.}
  \label{fig:dynkin}
\end{figure}

In the orthogonal basis $\left\{e_1,\dots,e_4\right\}$ the simple roots and the positive roots of $B_4$ are
\begin{equation}
  \label{eq:8}
 S_{B_4}= \{e_1 - e_2,\; e_2 - e_3,\; e_3 - e_4,\; e_4\}
\end{equation}
\begin{eqnarray*}
  \label{eq:19}
 \Delta^+_{B_4}=\left\{ (e_1 - e_2,\; e_2 - e_3,\; e_3 - e_4,\; e_4,\; e_1 - e_3,\; e_2 - e_4,\; e_3 + e_4,\; e_3,\; e_1 - e_4,\;\right.\\
 \left. e_2 + e_4,\; e_2,\; e_1 + e_4,\; e_2 + e_3,\; e_1,\; e_1 + e_3,\; e_1 + e_2\right\}
\end{eqnarray*}
Correspondingly for the embedded subalgebra $\mathfrak{a}=B_2$ we have
\begin{equation}
  \label{eq:26}
 S_{B_2}=\{e_3-e_4,e_4\}
\end{equation}
\begin{equation}
  \label{eq:27}
 \Delta^{+}_{\bot}= \left\{e_1-e_2,e_1+e_2,e_1,e_2\right\}
\end{equation}
and is the set of positive roots for the algebra $\mathfrak{a}_{\bot}=B_2$.
\begin{figure}[t]
  \centering
    \includegraphics[width=100mm,height=90mm]{B4_B2_anom_points.eps}
  \caption{Projected weights $-\frac{1}{s(\gamma_0)}\pi_{B_2}\left(\hat \Psi^{(0,1,0,2)}_{B_4}\right)$ with the dimensions of the corresponding $\mathfrak{a}_{\bot}$-modules.}
  \label{fig:B4B2anom}
\end{figure}

\begin{figure}[b]
  \centering
  \includegraphics[height=80mm]{B4_B2_fan.eps}
  \caption{The fan for $B_2\subset B_4$}
  \label{fig:B4B2Fan}
\end{figure}

@@

\begin{figure}[h!bt]
  \centering
  \includegraphics[width=135mm]{AffineA2_A1_fan.eps}
  \caption{The fan and $s(\gamma+\gamma_0)$ for $\hat{A_1}\longrightarrow \hat{A_2}$}
  \label{fig:AffineA2A1Fan}
\end{figure}

@@

\begin{figure}[h!tb]
%  \hspace*{-2cm}
  \includegraphics[width=140mm]{AffineA2_A1_anom.eps}
  \caption{The anomalous weights of the module $L^{(0,0;1;0)}_{\hat{A_2}}$}
  \label{fig:affine_A2_anom_point}
\end{figure}

The next step is to project the anomalous weights to $P_{\hat A_1}$.
The result up to the twelfth grade is presented in the Figure \ref{fig:AffineA2_A1_anom_proj}.
\begin{figure}[h!tb]
  \centering
  \includegraphics[width=130mm]{AffineA2_A1_proj_anom.eps}
  \caption{Projected anomalous weights of $L^{(0,0;1;0)}_{\hat{A_2}}$.}
  \label{fig:AffineA2_A1_anom_proj}
\end{figure}

Using the recurrent relation for the anomalous branching coefficients we get the result presented in Figure \ref{fig:AffineA2_A1_branching}. 
\begin{figure}[h!tb]
  \centering
  \includegraphics[width=130mm]{AffineA2_A1_branching.eps}
  \caption{Anomalous branching coefficients for $\hat{A_1}\subset \hat{A_2}$}
  \label{fig:AffineA2_A1_branching}
\end{figure}
There we see that inside the chamber $\bar{C}_{\hat{A}_1}$, which is isolated in the
Figure \ref{fig:AffineA2_A1_branching}, there are only two non-zero anomalous weights
both with multiplicities 1.
These are the highest weights of the $\mathfrak{a}$-submodules and their branching
coefficients. So the finite reducibility theorem holds and we get the decomposition
\begin{equation}
  \label{eq:43}
  L^{(0,0;1;0)}_{\hat{A_2}\downarrow \hat{A_1}}= L_{\hat{A_1}}^{(0;4;0)}\oplus L_{\hat{A_1}}^{(4;4;0)}.
\end{equation}

@@

Using the definition (\ref{fan-defined}) we get the fan
$\Gamma_{\hat{A_1} \longrightarrow  \hat{B_2} }$
with the corresponding values $s(\gamma+\gamma_0)$(see the Figure \ref{fig:AffineB2A1Fan}).
We restricted the computation to the twelfth grade.
\begin{figure}[h!bt]
  \centering
  \includegraphics[width=135mm]{AffineB2_A1_fan.eps}
  \caption{The fan for $\hat{A_1}\subset \hat{B_2}$}
  \label{fig:AffineB2A1Fan}
\end{figure}

@@

\begin{figure}[h!tb]
%  \hspace*{-2cm}
  \includegraphics[width=140mm]{AffineB2_A1_Anom.eps}
  \caption{The anomalous weights of $L^{(1,0;1;0)}_{\hat B_2 }$.
  The weights in the zero grade can be seen in the Figure \ref{fig:B2_A1}}.
  \label{fig:affine_B2_anom_point}
\end{figure}

Performing the next step of the algorithm \ref{sec:algorithm} we project the anomalous weights to
$P_{\hat{A_1}}$ and find the dimensions of the corresponding
$\mathfrak{a}_{\bot}$-modules $L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}$.
The result is presented in the Figure
\ref{fig:AffineB2_A1_anom_proj} up to the twelfth grade.
\ref{fig:AffineB2_A1_anom_proj}.
\begin{figure}[h!tb]
  \centering
  \includegraphics[width=120mm]{AffineB2_A1_proj_anom.eps}
  \caption{The projected anomalous weights $\pi_{\hat A_1}\left(\Psi^{(1,0;1;0)}_{\hat B_2}\right)$ and the dimensions of $\mathfrak{a}_{\bot}$-modules.}
  \label{fig:AffineB2_A1_anom_proj}
\end{figure}

@@

\begin{figure}[h!bt]
  \centering
  \includegraphics[width=120mm]{AffineB2_A1_branching.eps}
  \caption{Anomalous branching coefficients for $\hat{A_1}\subset \hat{B_2}$}
  \label{fig:AffineB2_A1_branching}
\end{figure}

>>>>

Such effects are due to the fact that the subalgebras $\mathfrak{a}$ and $\mathfrak{a}_{\bot}$ must not form a direct sum in $\mathfrak{g}$.  
Consider the case of such a regular embedding $\mathfrak{a}\subset \mathfrak{g}$ where both algebras are simple and the diagram of the subalgebra $\mathfrak{a}_{\bot}$ is not a subdiagram of the extended Dynkin diagram $\mathfrak{g}$.
 Drop the subdiagram of $\mathfrak{a}$ and the node $\alpha'$ that connects it with all the remaining nodes of the diagram of $\mathfrak{g}$. The remaining diagram is the diagram of the algebra $\mathfrak{\bar{a}}$ with $\mathrm{rank}(\mathfrak{\bar{a}}) = \mathrm{rank}(\mathfrak{g})-\mathrm{rank}(\mathfrak{a})$. It is clear that $\mathfrak{\bar{a}}\subset \mathfrak{a}_{\bot}$. So the question is whether $\mathfrak{a}_{\bot}$ has additional roots,  not the roots of $\mathfrak{\bar{a}}$ but the linear combinations of them. The answer is positive  when the set of angles between the roots of $\mathfrak{\bar{a}}$ does not contain all the angles between the roots of $\mathfrak{a}$. Then by reflecting the roots of $\mathfrak{a}$ by $s_{\alpha'}$ we get the additional roots of $\mathfrak{a}_{\bot}$. 
\begin{table}[ph]
\label{tab:diagrams}
\noindent  \centering{
\begin{tabular}{|l|l|l|l|}
  \hline
  $\mathfrak{g}$ & Extended diagram of $\mathfrak{g}$ & Diagrams of the subalgebras $\mathfrak{a},\; \mathfrak{a}_{\bot}$ \\
  \hline
  $A_n$ & \includegraphics{table1_1(l).eps} & \includegraphics{table1_1(r).eps} \\
  \hline                                                                     
  $B_n$ & \includegraphics{table1_2(l).eps} & \includegraphics{table1_2(r).eps} \\
  \hline                                                                     
  $C_n$ & \includegraphics{table1_3(l).eps} & \includegraphics{table1_3(r).eps} \\
  \hline                                                                     
  $D_n$ & \includegraphics{table1_4(l).eps} & \includegraphics{table1_4(r).eps} \\
  \hline
\end{tabular}
\caption{Subalgebras $\mathfrak{a},\; \mathfrak{a}_{\bot}$ for the classical series}}
\end{table}
If $\mathfrak{a}=B_{r_{\mathfrak{a}}}$  we see that $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\;\alpha_j,\; 1<i<j\leq r-r_{\mathfrak{a}}\}$ and $\mathfrak{a}_{\bot}=B_{r-r_{\mathfrak{a}}}$. This is the only case where the simple roots of $\mathfrak{a}_{\bot}$ can not be obtained from the extended Dynkin diagram, as can be seen in the Table \ref{tab:diagrams}.

@@

Let's describe the reduction of the fundamental representation of $L^{(1,0)=w_1}_{B_2}$ (the black vector in Figure \ref{fig:B2_A1}).
\begin{figure}[th]
  \noindent\centering{
    \includegraphics[width=80mm]{figure1.eps}
  }
  \caption{Regular embedding of $A_1$ into $B_2$}
  \label{fig:B2_A1}
\end{figure}
Circles indicate the weights of the singular element $\Psi^{(w_1)}$.
Now we are to factorise the Weyl group $W$ by $W_{\bot}=\left\{\omega_1\right\}$ and to construct the set $\left\{\omega(\mu+\rho)-\rho,\; \omega\in W_{\bot}\backslash W\right\}$. 
\begin{figure}[t]
  \noindent\centering{
    \includegraphics[width=80mm]{figure2.eps}
  }
  \caption{Anomalous weights and the corresponding $\mathfrak{a}_{\bot}=A_1$-modules for the embedding $A_1\subset B_2$}
  \label{fig:B2_A1_2}
\end{figure}
On the Figure \ref{fig:B2_A1_2} the weights of  the corresponding $\mathfrak{a}_{\bot}=A_1$-modules $L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}$ are indicated. Projecting them onto the root space of the subalgebra $\mathfrak{a}=A_1$ we get the anomalous weights with the corresponding multiplicities:
\begin{equation}
  \label{eq:25}
  (1,2),\; (0,-3),\; (-4,3),\; (-5,-2)
\end{equation}
For the set $\Gamma$ (using the definition (\ref{phi-d},\ref{fan-defined})) we have
\begin{equation}
  \label{eq:22}
  \Gamma_{A_1\subset B_2}=\left\{ (1,2),\; (2,-1) \right\}.
\end{equation}
Here the second component denotes the value of $s(\gamma)$. Applying this fan inside the $\bar{C}^{(0)}_{\mathfrak{a}}$ we get zeros for the weights greater than the first anomalous vector $(1)$, here $k^{(1,0)}_1=2$. For the last weight in $\bar{C}^{(0)}_{\mathfrak{a}}$ the formula (\ref{recurrent-relation}) gives
\begin{equation}
  \label{eq:23}
  k^{(1,0)}_{0}=-1\cdot k^{(1,0)}_2 +2\cdot k^{(1,0)}_1 - 3\cdot \delta_{0,0} = 1.
\end{equation}
The recurrence property defines the branching.

\subsection{Embedding of $B_2$ into $B_4$}
\label{sec:someth-high-dimens}
Consider the regular embedding $B_2 \longrightarrow B_4$.
The corresponding Dynkin diagrams are presented in the Figure \ref{fig:dynkin}.
\begin{figure}[h]
  \centering
  \includegraphics[width=50mm]{figure3.eps}
  \caption{The regular embedding $B_2\subset B_4$ described by dropping the node.}
  \label{fig:dynkin}
\end{figure}

In the orthogonal basis $\left\{e_1,\dots,e_4\right\}$ the simple roots and the positive roots of $B_4$ are
\begin{equation}
  \label{eq:8}
 S_{B_4}= \{e_1 - e_2,\; e_2 - e_3,\; e_3 - e_4,\; e_4\}
\end{equation}
\begin{eqnarray*}
  \label{eq:19}
 \Delta^+_{B_4}=\left\{ (e_1 - e_2,\; e_2 - e_3,\; e_3 - e_4,\; e_4,\; e_1 - e_3,\; e_2 - e_4,\; e_3 + e_4,\; e_3,\; e_1 - e_4,\;\right.\\
 \left. e_2 + e_4,\; e_2,\; e_1 + e_4,\; e_2 + e_3,\; e_1,\; e_1 + e_3,\; e_1 + e_2\right\}
\end{eqnarray*}
Correspondingly for the embedded subalgebra $\mathfrak{a}=B_2$ we have
\begin{equation}
  \label{eq:26}
 S_{B_2}=\{e_3-e_4,e_4\}
\end{equation}
\begin{equation}
  \label{eq:27}
 \Delta^{+}_{\bot}= \left\{e_1-e_2,e_1+e_2,e_1,e_2\right\}
\end{equation}
and is the set of positive roots for the algebra $\mathfrak{a}_{\bot}=B_2$.
\begin{figure}[t]
  \centering
    \includegraphics[width=100mm,height=90mm]{figure4.eps}
  \caption{Projected weights $-\frac{1}{s(\gamma_0)}\pi_{B_2}\left(\hat \Psi^{(0,1,0,2)}_{B_4}\right)$ with the dimensions of the corresponding $\mathfrak{a}_{\bot}$-modules.}
  \label{fig:B4B2anom}
\end{figure}

\begin{figure}[b]
  \centering
  \includegraphics[height=80mm]{figure5.eps}
  \caption{The fan for $B_2\subset B_4$}
  \label{fig:B4B2Fan}
\end{figure}

@@

\begin{figure}[h!bt]
  \centering
  \includegraphics[width=135mm]{figure6.eps}
  \caption{The fan and $s(\gamma+\gamma_0)$ for $\hat{A_1}\longrightarrow \hat{A_2}$}
  \label{fig:AffineA2A1Fan}
\end{figure}

@@

\begin{figure}[h!tb]
%  \hspace*{-2cm}
  \includegraphics[width=140mm]{figure7.eps}
  \caption{The anomalous weights of the module $L^{(0,0;1;0)}_{\hat{A_2}}$}
  \label{fig:affine_A2_anom_point}
\end{figure}

The next step is to project the anomalous weights to $P_{\hat A_1}$.
The result up to the twelfth grade is presented in the Figure \ref{fig:AffineA2_A1_anom_proj}.
\begin{figure}[h!tb]
  \centering
  \includegraphics[width=130mm]{figure8.eps}
  \caption{Projected anomalous weights of $L^{(0,0;1;0)}_{\hat{A_2}}$.}
  \label{fig:AffineA2_A1_anom_proj}
\end{figure}

Using the recurrent relation for the anomalous branching coefficients we get the result presented in Figure \ref{fig:AffineA2_A1_branching}. 
\begin{figure}[h!tb]
  \centering
  \includegraphics[width=130mm]{figure9.eps}
  \caption{Anomalous branching coefficients for $\hat{A_1}\subset \hat{A_2}$}
  \label{fig:AffineA2_A1_branching}
\end{figure}
There we see that inside the chamber $\bar{C}_{\hat{A}_1}$, which is isolated in the
Figure \ref{fig:AffineA2_A1_branching}, there are only two non-zero anomalous weights
both with multiplicities 1.
These are the highest weights of the $\mathfrak{a}$-submodules and their branching
coefficients. So the finite reducibility theorem holds and we get the decomposition
\begin{equation}
  \label{eq:43}
  L^{(0,0;1;0)}_{\hat{A_2}\downarrow \hat{A_1}}= L_{\hat{A_1}}^{(0;4;0)}\oplus L_{\hat{A_1}}^{(4;4;0)}.
\end{equation}

@@

Using the definition (\ref{fan-defined}) we get the fan
$\Gamma_{\hat{A_1} \longrightarrow  \hat{B_2} }$
with the corresponding values $s(\gamma+\gamma_0)$(see the Figure \ref{fig:AffineB2A1Fan}).
We restricted the computation to the twelfth grade.
\begin{figure}[h!bt]
  \centering
  \includegraphics[width=135mm]{figure10.eps}
  \caption{The fan for $\hat{A_1}\subset \hat{B_2}$}
  \label{fig:AffineB2A1Fan}
\end{figure}

@@

\begin{figure}[h!tb]
%  \hspace*{-2cm}
  \includegraphics[width=140mm]{figure11.eps}
  \caption{The anomalous weights of $L^{(1,0;1;0)}_{\hat B_2 }$.
  The weights in the zero grade can be seen in the Figure \ref{fig:B2_A1}}.
  \label{fig:affine_B2_anom_point}
\end{figure}

Performing the next step of the algorithm \ref{sec:algorithm} we project the anomalous weights to
$P_{\hat{A_1}}$ and find the dimensions of the corresponding
$\mathfrak{a}_{\bot}$-modules $L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}$.
The result is presented in the Figure
\ref{fig:AffineB2_A1_anom_proj} up to the twelfth grade.
\ref{fig:AffineB2_A1_anom_proj}.
\begin{figure}[h!tb]
  \centering
  \includegraphics[width=120mm]{figure12.eps}
  \caption{The projected anomalous weights $\pi_{\hat A_1}\left(\Psi^{(1,0;1;0)}_{\hat B_2}\right)$ and the dimensions of $\mathfrak{a}_{\bot}$-modules.}
  \label{fig:AffineB2_A1_anom_proj}
\end{figure}

@@

\begin{figure}[h!bt]
  \centering
  \includegraphics[width=120mm]{figure13.eps}
  \caption{Anomalous branching coefficients for $\hat{A_1}\subset \hat{B_2}$}
  \label{fig:AffineB2_A1_branching}
\end{figure}
